---
title: "SAD space report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(drake)
library(sadspace)
library(dplyr)
library(ggplot2)
library(ggpmisc)
```

```{r load stuff}

sv <- define_statevars() %>%
  assign_ptable() %>%
  dplyr::filter(p_table != "none")

#di_df <- read.csv(here::here("analysis", "di_dfs.csv"), stringsAsFactors = F)
fs_size_dat <- read.csv(here::here("analysis", "fs_size_dat.csv"), stringsAsFactors = F)
di_summary_df <- read.csv(here::here("analysis", "di_summary_df.csv"), stringsAsFactors = F)

```

Here is the range of S and N space covered:

```{r plot sn space, fig.dim = c(3.5, 3.5)}

ggplot(sv, aes(s0, n0, color = n0/s0)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "plasma") +
  theme(legend.position = "top")

ggplot(sv, aes(log(s0), log(n0), color = log(n0/s0))) +
  geom_point() +
  theme_bw()+
  scale_color_viridis_c(option = "plasma")+
  theme(legend.position = "top")
```

Things are generally easier to see on the log axis, so let's stick with that.

## Size and unique elements from FS

Here is how the size of the feasible set varies with S, N, and N/S:

```{r plot size of fs 1}
# 
# fs_size_dat <- di_df %>%
#   select(s0, n0, nparts, nunique) %>%
#   distinct() %>%
#   mutate(log_nparts = log(as.numeric(nparts)),
#          log_nunique = log(nunique))

ggplot(fs_size_dat, aes(log(s0), log(n0), color =log_nparts)) +
  geom_point() + 
  theme_bw()+
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8)+
  theme(legend.position = "top")

```


This is *on a log scale*, so up in that right corner is *1.942426e+130*. 

For now I am interested in lognparts up to about 20 

```{r nparts up to 12}


ggplot(filter(fs_size_dat, log_nparts <= 20), aes(log(s0), log(n0), color =log_nparts)) +
  geom_point() + 
  theme_bw()+
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8)+
  theme(legend.position = "top")

fs_size_dat <- fs_size_dat %>%
  mutate(more_than_11 = log_nparts >= 11)


ggplot(filter(fs_size_dat, log_nparts <= 20), aes(log(s0), log(n0), color =more_than_11)) +
  geom_point() + 
  theme_bw()+
  scale_color_viridis_d(option = "plasma", begin = .1, end = .8)+
  theme(legend.position = "top")

```


These are the communities we'll pull out:


```{r plot focal communities}


focal_communities <- data.frame(
  s0 = c(3, 3, 3, 5, 5, 5,  21, 21, 149, 149),
  n0 = c(13, 149, 1097, 13, 149, 1097, 34, 91, 156, 245)
)


focal_communities <- focal_communities %>%
  mutate(target_name = paste0("fs_", s0, "_", n0))

## this is scrap code to pull 3000 unique sims per focal community from cache on hpg. Drawing more gives v large .csvs for large communities.
# for(i in 1:nrow(focal_communities)) {
# thisobj = get_object(focal_communities$target_name[i])
# thisobj <- thisobj %>% filter(sim %in% unique(sim)[1:3000])
# write.csv(thisobj, here::here("analysis", "reports", paste0(focal_communities$target_name[i], ".csv")), row.names = F)
# }
focal_communities <- left_join(focal_communities, fs_size_dat, by = c("s0", "n0"))

ggplot(filter(fs_size_dat, log_nparts <= 20), aes(log(s0), log(n0), color =more_than_11)) +
  geom_point(alpha = .2) + 
  geom_point(data = focal_communities, alpha = 1, shape = 8, size = 2) +
  theme_bw()+
  scale_color_viridis_d(option = "plasma", begin = .1, end = .8)+
  theme(legend.position = "top")

focal_communities <- focal_communities %>%
  mutate(file_name = here::here("analysis", "reports", paste0(target_name, ".csv")))

focal_fs <- list()
for(i in 1:nrow(focal_communities)) {
  focal_fs[[i]] <- read.csv(focal_communities$file_name[i], stringsAsFactors = F)
  
  focal_fs[[i]] <-  focal_fs[[i]] %>%
    group_by(sim, nparts, s0, n0) %>%
    mutate(skew = e1071::skewness(abund),
           simpson = vegan::diversity(abund, index = "simpson")) %>%
    ungroup() %>%
    mutate(s0_f = as.factor(s0),
           n0_f = as.factor(n0))
}

focal_fs <- bind_rows(focal_fs)

focal_fs <- mutate(focal_fs, more_than_11 = log(as.numeric(nparts)) >= 11)

focal_fs_summary <- focal_fs %>%
  select(-rank, -abund, -s0_f, -n0_f) %>%
  distinct() %>%
  mutate(community = paste0(s0, "_", n0))
``` 

Here are the FS of those distributions:

```{r all fs}
ggplot(focal_fs, aes(x = rank, y = abund, group = sim, color = more_than_11)) +
  geom_line(alpha = .2) +
  facet_wrap(vars(s0_f, n0_f), scales = "free") +
  theme_bw()

```

Here are histograms of skew/even for these distributions:

```{r hists}
ggplot(focal_fs_summary, aes(x = skew, fill = more_than_11)) +
  geom_histogram() +
  theme_bw() +
  facet_grid(rows = vars(n0), cols = vars(s0), scales = "free")

ggplot(focal_fs_summary, aes(x = simpson, fill = more_than_11)) +
  geom_histogram() +
  theme_bw() +
  facet_grid(rows = vars(n0), cols = vars(s0), scales = "free")
```


